Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-10-11 paper_comparison/paper_comparison/paper_comparison.Rmd
fix some references

As summarised in Table \ref{tab:model-factor-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous. Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ were used in experiment I and II respectively. Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters. A large value of $\sigma$ will increase the variation of the error of the polynomial model and decrease the visibility of the visual pattern. The parameter $b$ controls the ratio of the standard deviation of the heteroskedasticity across the domain of the regressor. Given $x \neq a$, a larger value of $b$ will lead to a larger ratio of the variance at $x$ to the variance at $x - a = 0$, making the visual pattern more obvious.

@@

```{r different-sigma, fig.height = 2, fig.cap="Example residual plots of the linear regression model used in experiment I. Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)

@@

```{r different-b, fig.height = 2, fig.cap="Example residual plots of the linear regression model used in experiment II. Five different values of $b$ are included in the experiment to control the strength of the signal."}
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

@@

Visual inference and conventional hypothesis testing are two distinct statistical procedures. As shown in \ref{fig:p-value-comparison}, the visual test almost always gives higher p-values than the conventional test, suggesting that humans have lower confidence and much higher tolerance of the residual departures than the numerical procedure. In fact, only around 58% of the cases, both tests will agree on the rejection decision given 95% confidence level. And in 41% of the cases, visual test does not reject the null hypothesis but conventional test does. Although we know that the all the designed lineups contain various degrees of residual departures and the conventional test is telling truth, analysts and general public as the consumers of the output will possibly not believe in or at least not response to the "false alarm" if they can not see the departures directly from the data plot and view it as impactful. The sensitivity of the conventional test may distract and discourage us from finding a good linear approximation to the data we try to model. This also partly explains why abundance of literature suggesting the use of data plot in regression diagnostics.

```{r eval = FALSE}
vi_survey %>%
  filter(lineup_id <= 576) %>%
  filter(is.na(b) | b != 0) %>%
  mutate(test = p_value > 0.05 & conventional_p_value < 0.05) %>%
  {mean(.$test)}
```

```{r p-value-comparison, fig.cap="..."}
vi_survey %>%
  filter(lineup_id <= 576) %>%
  filter(is.na(b) | b != 0) %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  # filter(x_dist == "uniform") %>%
  ggplot() +
  geom_point(aes(p_value, conventional_p_value), alpha = 0.4) +
  geom_abline() +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(~type)
```

@@

Figure \ref{fig:power-poly-uniform} shows the estimated power of visual test against natural logarithm of the effect with comparison to the power of an exact test - F-test, and the power of two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions. In overall, the power of all four tests increases as the effect becomes larger. The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while others respond inactively to the change of effect and remain lower than 25% throughout the period, showing that as an exact test, the F-test is relatively more sensitive to the type of model defects that being considered. The power of visual test arises steadily and nearly linearly to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features. Other two inappropriate conventional tests shows improvement at the same time but at a lower rate. This coincides the point made by @cook1982residuals mentioned in \ref{hypothesis-testing} that residual-based tests for a specific type of model defect are sensitive to other types of model defects. At $log_e(E) = 6$, the power curve of F-test reaches almost 100% followed by the visual test by a small margin. The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 30% respectively.

What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test. The margin is largest at around $log_e(E) = 2$. An example lineup is included in Figure \ref{fig:power-overview} where none of subjects detect the actual data plot positioned at panel 14. It demonstrates that at this level of difficulty, the designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model. From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public even though it is true. At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger. At this level of difficulty, the designed visual features are usually detectable but it may not stand out from the lineup as other null plots may happen to include outliers or visual patterns that are are considered to be more attractive by human, and thus recognized as the most different plot. Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison. The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 17, where two out of five subjects detect it. It can be observed that a "U-shape" is presented in plot 17, but the signal is not strong enough to attract all five subjects, resulting in a visual p-value sightly above the desired significance level $\alpha = 0.05$. At $log_e(E) = 4$ and $log_e(E) = 5$, the designed visual features become much clear and attractive, leading to a high percentage of rejection of the null hypothesis. Figure \ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It is also interesting to observe that the power curve of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible.

@@

```{r fig:power-poly-uniform, fig.cap = "..."}
vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) + geom_point(data = ~filter(.x, unique_lineup_id %in% c("poly_502", "poly_78", "poly_6", "poly_447")), aes(log_effect_size, prop_detect), alpha = 1, col = "red", size = 2) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual p-value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                filter(x_dist == "uniform") %>%
                pivot_longer(F_p_value:SW_p_value), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  theme_light(base_size = 5) -> p

@@

The impact of the distribution of $X_raw$ on the power is shown in Figure \ref{fig:dist-power}. The power curve of F-test is stable across different distributions, while the visual test has a steeper power curve for normal and uniform distribution. BP-test performs worse for discrete uniform distribution and uniform distribution but has relatively high power for normal distribution. SW-test outperforms BP-test for discrete uniform distribution but remains as the worst test for other distributions. The results indicate those inappropriate residual-based tests are sensitive to the distribution of the regressor.

>>>>

As summarised in Table \ref{tab:model-factor-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous. Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ were used in experiment I and II respectively. Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters. A large value of $\sigma$ will increase the variation of the error of the non-linear model and decrease the visibility of the visual pattern. The parameter $b$ controls the ratio of the standard deviation of the heteroskedasticity across the domain of the regressor. Given $x \neq a$, a larger value of $b$ will lead to a larger ratio of the variance at $x$ to the variance at $x - a = 0$, making the visual pattern more obvious.

@@

```{r different-sigma, fig.height = 2, fig.cap="Four different values of $\\sigma$ are used in the experiment I to control the strength of the signal."}
ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)

@@

```{r different-b, fig.height = 2, fig.cap="Five different values of $b$ are used in experiment II to control the strength of the signal."}
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

@@

Visual inference and conventional hypothesis testing are two distinct statistical procedures. As displayed in Figure \ref{fig:p-value-comparison}, the visual test almost always gives higher p-values than the conventional test in both polynomial and heteroskedasticity models, suggesting that humans have lower confidence and much higher tolerance of the residual departures than the numerical procedure. In fact, only around 58% of the cases, both tests will agree on the rejection decision given 95% confidence level. And in 41% of the cases, visual test does not reject the null hypothesis but conventional test does. Although we know that the all the designed lineups contain various degrees of residual departures and the conventional test is telling truth, analysts and general public as the consumers of the output will possibly not believe in or at least not response to the "false alarm" if they can not see the departures directly from the data plot and view it as impactful. The sensitivity of the conventional test may distract and discourage us from finding a good linear approximation to the data we try to model. This also partly explains why abundance of literature suggesting the use of data plot in regression diagnostics.

```{r eval = FALSE}
vi_survey %>%
  filter(lineup_id <= 576) %>%
  filter(is.na(b) | b != 0) %>%
  mutate(test = p_value > 0.05 & conventional_p_value < 0.05) %>%
  {mean(.$test)}

vi_survey %>%
  filter(lineup_id <= 576) %>%
  filter(is.na(b) | b != 0) %>%
ggplot() +
  geom_histogram(aes(p_value), bins = 300) +
  scale_x_sqrt() +
 facet_wrap(~type)
```

```{r p-value-comparison, fig.height=4, fig.cap="..."}
vi_survey %>%
  filter(lineup_id <= 576) %>%
  filter(is.na(b) | b != 0) %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  # filter(x_dist == "uniform") %>%
  ggplot() +
  geom_point(aes(p_value, conventional_p_value), alpha = 0.4) +
  geom_abline() +
  scale_x_sqrt() +
  scale_y_sqrt() +
  facet_wrap(~type)
```

@@

Figure \ref{fig:power-poly-uniform} shows the estimated power of visual test on lineups with uniform distribution $X_{raw}$ against natural logarithm of the effect with comparison to the power of an exact test - F-test, and the power of two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions. In overall, the power of all four tests increases as the effect becomes larger. The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while others respond inactively to the change of effect and remain lower than 25% throughout the period, showing that as an exact test, the F-test is relatively more sensitive to the type of model defects that being considered. The power of visual test arises steadily and nearly linearly to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features. Other two inappropriate conventional tests shows improvement at the same time but at a lower rate. This coincides the point made by @cook1982residuals mentioned in \ref{hypothesis-testing} that residual-based tests for a specific type of model defect are sensitive to other types of model defects. At $log_e(E) \approx 6$, the power curve of F-test reaches almost 100% followed by the visual test by a small margin. The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 30% respectively.

What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test. The margin is largest at around $log_e(E) = 2$. An example lineup is included in Figure \ref{fig:power-poly-uniform} where none of subjects detect the actual data plot positioned at panel 10. It demonstrates that at this level of difficulty, the designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model. From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public even though it is true. At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger. At this level of difficulty, the designed visual features are usually detectable but it may not stand out from the lineup as other null plots may happen to include outliers or visual patterns that are are considered to be more attractive by human, and thus recognized as the most different plot. Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison. The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 7, where four of 11 subjects detect it. It can be observed that a "U-shape" is presented in plot 7, but the signal is not strong enough to attract all five subjects, resulting in a visual p-value sightly above the desired significance level $\alpha = 0.05$. At $log_e(E) = 4$, the designed visual features become much clear and attractive, leading to a high percentage of rejection of the null hypothesis. Figure \ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It is also interesting to observe that the power curve of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible.

@@

```{r power-poly-uniform, fig.cap = "..."}
vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) + geom_point(data = ~filter(.x, unique_lineup_id %in% c("poly_502", "poly_78", "poly_6", "poly_447")), aes(log_effect_size, prop_detect), alpha = 1, col = "red", size = 2) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual p-value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                filter(x_dist == "uniform") %>%
                pivot_longer(F_p_value:SW_p_value), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  theme_light(base_size = 5) -> p

@@

The impact of the distribution of $X_raw$ on the power is displayed in Figure \ref{fig:dist-power}. The power curve of F-test is stable across different distributions, while the visual test has a steeper power curve for normal and uniform distribution. BP-test performs worse for discrete uniform distribution and uniform distribution but has relatively high power for normal distribution. SW-test outperforms BP-test for discrete uniform distribution but remains as the worst test for other distributions. The results indicate those inappropriate residual-based tests are sensitive to the distribution of the regressor.

### 2022-10-11 paper_comparison/paper_comparison/paper_comparison.tex
fix some references

As summarised in Table \ref{tab:model-factor-table}, three additional
parameters \(n\), \(\sigma\) and \(b\) were used to control the strength
of the signal so that different difficulty levels of lineups were
generated, and therefore, the estimated power curve would be smooth and
continuous. Parameter \(\sigma \in \{0.5, 1, 2, 4\}\) and
\(b \in \{0.25, 1, 4, 16, 64\}\) were used in experiment I and II
respectively. Figure \ref{fig:different-sigma} and \ref{fig:different-b}
demonstrate the impact of these two parameters. A large value of
\(\sigma\) will increase the variation of the error of the polynomial
model and decrease the visibility of the visual pattern. The parameter
\(b\) controls the ratio of the standard deviation of the
heteroskedasticity across the domain of the regressor. Given
\(x \neq a\), a larger value of \(b\) will lead to a larger ratio of the
variance at \(x\) to the variance at \(x - a = 0\), making the visual
pattern more obvious.

@@

\caption{Example residual plots of the linear regression model used in experiment I. Four different values of $\sigma$ are included in the experiment to control the strength of the signal.}\label{fig:different-sigma}
\end{figure}

@@

\caption{Example residual plots of the linear regression model used in experiment II. Five different values of $b$ are included in the experiment to control the strength of the signal.}\label{fig:different-b}
\end{figure}

@@

Visual inference and conventional hypothesis testing are two distinct
statistical procedures. As shown in \ref{fig:p-value-comparison}, the
visual test almost always gives higher p-values than the conventional
test, suggesting that humans have lower confidence and much higher
tolerance of the residual departures than the numerical procedure. In
fact, only around 58\% of the cases, both tests will agree on the
rejection decision given 95\% confidence level. And in 41\% of the
cases, visual test does not reject the null hypothesis but conventional
test does. Although we know that the all the designed lineups contain
various degrees of residual departures and the conventional test is
telling truth, analysts and general public as the consumers of the
output will possibly not believe in or at least not response to the
``false alarm'' if they can not see the departures directly from the
data plot and view it as impactful. The sensitivity of the conventional
test may distract and discourage us from finding a good linear
approximation to the data we try to model. This also partly explains why
abundance of literature suggesting the use of data plot in regression
diagnostics.

@@

Figure \ref{fig:power-poly-uniform} shows the estimated power of visual
test against natural logarithm of the effect with comparison to the
power of an exact test - F-test, and the power of two other
residual-based conventional tests commonly used in regression
diagnostics but for testing other departures from the model assumptions.
In overall, the power of all four tests increases as the effect becomes
larger. The power curve of F-test climbs aggressively from 25\% to
around 90\% as \(log_e(E)\) increases from 0 to 2, while others respond
inactively to the change of effect and remain lower than 25\% throughout
the period, showing that as an exact test, the F-test is relatively more
sensitive to the type of model defects that being considered. The power
of visual test arises steadily and nearly linearly to around 90\% as
\(log_e(E)\) increases from 2 to 5, suggesting that the effect starts to
make noticeable impact on the degree of the presence of the designed
visual features. Other two inappropriate conventional tests shows
improvement at the same time but at a lower rate. This coincides the
point made by \citet{cook1982residuals} mentioned in
\ref{hypothesis-testing} that residual-based tests for a specific type
of model defect are sensitive to other types of model defects. At
\(log_e(E) = 6\), the power curve of F-test reaches almost 100\%
followed by the visual test by a small margin. The power of
Breusch--Pagan test and Shapiro--Wilk test reach around 75\% and 30\%
respectively.

What truly impress us is the huge difference between the estimated power
of visual test and the estimated power of F-test. The margin is largest
at around \(log_e(E) = 2\). An example lineup is included in Figure
\ref{fig:power-overview} where none of subjects detect the actual data
plot positioned at panel 14. It demonstrates that at this level of
difficulty, the designed visual feature is rarely visible, making the
actual data plot indistinguishable from residual plots simulated from
the assumed model. From a communication perspective, given the fact that
the visual difference is unperceivable, the argument that non-linearity
present in the fitted model is less convincing to the public even though
it is true. At around \(log_e(E) = 3\), the margin gets smaller as the
chance of identifying the actual data plot becomes larger. At this level
of difficulty, the designed visual features are usually detectable but
it may not stand out from the lineup as other null plots may happen to
include outliers or visual patterns that are are considered to be more
attractive by human, and thus recognized as the most different plot.
Without knowing the designed visual features beforehand, it is actually
hard to identify the actual data plot by pure image comparison. The
corresponding example lineup for \(log_e(E) = 3\) shown in Figure
\ref{fig:fig:power-poly-uniform} has the actual data plot positioned at
panel 17, where two out of five subjects detect it. It can be observed
that a ``U-shape'' is presented in plot 17, but the signal is not strong
enough to attract all five subjects, resulting in a visual p-value
sightly above the desired significance level \(\alpha = 0.05\). At
\(log_e(E) = 4\) and \(log_e(E) = 5\), the designed visual features
become much clear and attractive, leading to a high percentage of
rejection of the null hypothesis. Figure
\ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It
is also interesting to observe that the power curve of visual test
behaves similarly to those inappropriate conventional tests until the
designed visual features really becomes visible.

@@

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/fig:power-poly-uniform-1}

@@

\caption{...}\label{fig:fig:power-poly-uniform}
\end{figure}

@@

The impact of the distribution of \(X_raw\) on the power is shown in
Figure \ref{fig:dist-power}. The power curve of F-test is stable across
different distributions, while the visual test has a steeper power curve
for normal and uniform distribution. BP-test performs worse for discrete
uniform distribution and uniform distribution but has relatively high
power for normal distribution. SW-test outperforms BP-test for discrete
uniform distribution but remains as the worst test for other
distributions. The results indicate those inappropriate residual-based
tests are sensitive to the distribution of the regressor.

>>>>

As summarised in Table \ref{tab:model-factor-table}, three additional
parameters \(n\), \(\sigma\) and \(b\) were used to control the strength
of the signal so that different difficulty levels of lineups were
generated, and therefore, the estimated power curve would be smooth and
continuous. Parameter \(\sigma \in \{0.5, 1, 2, 4\}\) and
\(b \in \{0.25, 1, 4, 16, 64\}\) were used in experiment I and II
respectively. Figure \ref{fig:different-sigma} and \ref{fig:different-b}
demonstrate the impact of these two parameters. A large value of
\(\sigma\) will increase the variation of the error of the non-linear
model and decrease the visibility of the visual pattern. The parameter
\(b\) controls the ratio of the standard deviation of the
heteroskedasticity across the domain of the regressor. Given
\(x \neq a\), a larger value of \(b\) will lead to a larger ratio of the
variance at \(x\) to the variance at \(x - a = 0\), making the visual
pattern more obvious.

@@

\caption{Four different values of $\sigma$ are used in the experiment I to control the strength of the signal.}\label{fig:different-sigma}
\end{figure}

@@

\caption{Five different values of $b$ are used in experiment II to control the strength of the signal.}\label{fig:different-b}
\end{figure}

@@

Visual inference and conventional hypothesis testing are two distinct
statistical procedures. As displayed in Figure
\ref{fig:p-value-comparison}, the visual test almost always gives higher
p-values than the conventional test in both polynomial and
heteroskedasticity models, suggesting that humans have lower confidence
and much higher tolerance of the residual departures than the numerical
procedure. In fact, only around 58\% of the cases, both tests will agree
on the rejection decision given 95\% confidence level. And in 41\% of
the cases, visual test does not reject the null hypothesis but
conventional test does. Although we know that the all the designed
lineups contain various degrees of residual departures and the
conventional test is telling truth, analysts and general public as the
consumers of the output will possibly not believe in or at least not
response to the ``false alarm'' if they can not see the departures
directly from the data plot and view it as impactful. The sensitivity of
the conventional test may distract and discourage us from finding a good
linear approximation to the data we try to model. This also partly
explains why abundance of literature suggesting the use of data plot in
regression diagnostics.

@@

Figure \ref{fig:power-poly-uniform} shows the estimated power of visual
test on lineups with uniform distribution \(X_{raw}\) against natural
logarithm of the effect with comparison to the power of an exact test -
F-test, and the power of two other residual-based conventional tests
commonly used in regression diagnostics but for testing other departures
from the model assumptions. In overall, the power of all four tests
increases as the effect becomes larger. The power curve of F-test climbs
aggressively from 25\% to around 90\% as \(log_e(E)\) increases from 0
to 2, while others respond inactively to the change of effect and remain
lower than 25\% throughout the period, showing that as an exact test,
the F-test is relatively more sensitive to the type of model defects
that being considered. The power of visual test arises steadily and
nearly linearly to around 90\% as \(log_e(E)\) increases from 2 to 5,
suggesting that the effect starts to make noticeable impact on the
degree of the presence of the designed visual features. Other two
inappropriate conventional tests shows improvement at the same time but
at a lower rate. This coincides the point made by
\citet{cook1982residuals} mentioned in \ref{hypothesis-testing} that
residual-based tests for a specific type of model defect are sensitive
to other types of model defects. At \(log_e(E) \approx 6\), the power
curve of F-test reaches almost 100\% followed by the visual test by a
small margin. The power of Breusch--Pagan test and Shapiro--Wilk test
reach around 75\% and 30\% respectively.

What truly impress us is the huge difference between the estimated power
of visual test and the estimated power of F-test. The margin is largest
at around \(log_e(E) = 2\). An example lineup is included in Figure
\ref{fig:power-poly-uniform} where none of subjects detect the actual
data plot positioned at panel 10. It demonstrates that at this level of
difficulty, the designed visual feature is rarely visible, making the
actual data plot indistinguishable from residual plots simulated from
the assumed model. From a communication perspective, given the fact that
the visual difference is unperceivable, the argument that non-linearity
present in the fitted model is less convincing to the public even though
it is true. At around \(log_e(E) = 3\), the margin gets smaller as the
chance of identifying the actual data plot becomes larger. At this level
of difficulty, the designed visual features are usually detectable but
it may not stand out from the lineup as other null plots may happen to
include outliers or visual patterns that are are considered to be more
attractive by human, and thus recognized as the most different plot.
Without knowing the designed visual features beforehand, it is actually
hard to identify the actual data plot by pure image comparison. The
corresponding example lineup for \(log_e(E) = 3\) shown in Figure
\ref{fig:fig:power-poly-uniform} has the actual data plot positioned at
panel 7, where four of 11 subjects detect it. It can be observed that a
``U-shape'' is presented in plot 7, but the signal is not strong enough
to attract all five subjects, resulting in a visual p-value sightly
above the desired significance level \(\alpha = 0.05\). At
\(log_e(E) = 4\), the designed visual features become much clear and
attractive, leading to a high percentage of rejection of the null
hypothesis. Figure \ref{fig:fig:power-poly-uniform} gives example
lineups of such cases. It is also interesting to observe that the power
curve of visual test behaves similarly to those inappropriate
conventional tests until the designed visual features really becomes
visible.

@@

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/power-poly-uniform-1}

@@

\caption{...}\label{fig:power-poly-uniform}
\end{figure}

@@

The impact of the distribution of \(X_raw\) on the power is displayed in
Figure \ref{fig:dist-power}. The power curve of F-test is stable across
different distributions, while the visual test has a steeper power curve
for normal and uniform distribution. BP-test performs worse for discrete
uniform distribution and uniform distribution but has relatively high
power for normal distribution. SW-test outperforms BP-test for discrete
uniform distribution but remains as the worst test for other
distributions. The results indicate those inappropriate residual-based
tests are sensitive to the distribution of the regressor.
